---
title: "Online Retail Customer Retention Analysis"
author: "Alex"
date: "2025-04-09"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

# Introduction
This report analyses customer purchasing patterns to improve retention for an online retail company

# Data Loading and Cleaning
```{r}
# Load the tidyverse package
library(tidyverse)
library(readr)

# Load the dataset
retail_data <- read_csv("OnlineRetail.csv")
```
```{r}
# Remove rows with missing CustomerID and negative quantities
clean_data <- retail_data %>%
  filter(!is.na(CustomerID), Quantity > 0)

# Convert InvoiceDate to a date format
clean_data <- clean_data %>%
  mutate(InvoiceDate = as.Date(InvoiceDate, format = "%m/%d/%Y %H:%M"))

# Look at the first few rows
head(clean_data)
```

# Analysis
Let’s explore customer purchasing patterns.
```{r}
# Count transactions per customer
customer_frequency <- clean_data %>%
  group_by(CustomerID) %>%
  summarise(Transactions = n()) %>%
  arrange(desc(Transactions))

# Show top 5 customers
head(customer_frequency, 5)
```

```{r}
# Plot a histogram of transaction counts
ggplot(customer_frequency, aes(x = Transactions)) +
  geom_histogram(binwidth = 1, fill = "blue", color = "black") +
  labs(title = "Distribution of Customer Purchase Frequency",
       x = "Number of Transactions", y = "Count of Customers")
```
```{r}
# Calculate total spend per transaction and average per customer
clean_data <- clean_data %>%
  mutate(TotalSpend = Quantity * UnitPrice)

avg_spend <- clean_data %>%
  group_by(CustomerID) %>%
  summarise(AvgSpend = mean(TotalSpend)) %>%
  arrange(desc(AvgSpend))

head(avg_spend, 5)
```

# Findings
Most customers buy only once, but a small group of repeat buyers drives sales. The top spenders average higher transaction values.

# Recommendations
- Offer discounts to one-time buyers to encourage repeat purchases.
- Reward top spenders with loyalty perks.
